---
title: "ASN5.2. Modeling for Inference"
author: "2023 FIRE298 Sustainability Analytics"
format: html
editor: visual
---

In this assignment, you will use the difference-in-difference model framework to estimate the impact of the treatment variable.

### **Part 1. Explaining the Data**

Instructions: Run the script below to download and declare the required packages.

```{r}
#| message: false
#install.packages("tidyverse")
library("tidyverse")
```

Instructions: Upload your team's final dataframe into Posit Cloud. Use the read.csv function to upload the dataframe into the Posit Cloud Environment.

```{r}
panel <- read.csv("paneldata.csv")
```

**Question 1: Which variable indicates the treatment *group*? If your dataset does not have one yet, please create it using the mutate and ifelse function (2 points)**

Answer: The treatment group is when there is an incident of a crime, which is when the variable equals one. Indicating there is pollution.

*Feedback: The treatment variable is treatment, which indicates whether wind blows from the Wheelabrator Incinerator to each neighborhood or not. (1 point deduction)*

**Question 2: Which variable indicates the treatment *period?* If your dataset does not have one yet, please create it using the mutate and ifelse function (2 points)**

Answer: The treatment period is when there is an incident of a crime, which is when the variable equals one. Indicating there is pollution.

**Question 3: Run a DiD regression with the treatment group** $d_i$**, treatment period** $t$**, and its interaction** $d_{it} \times t$ **to estimate their impact on the outcome variable. This regression can be represent by the equation** $Y_{it}=\beta_0+\beta_1 d_i+\beta_2 t + \gamma (d_{it}\times t) + \epsilon_{it}$

**Afterward, print the regression results by using the function summary(*ModelName*). An example of this can be found in the lecture slides 2023 FIRE298 Week 12 Part 2 Slide 5.**

```{r}
summary(mode <- lm(total_violent_crime~treatment, data = panel))
```

**Based on the regression, what is the value of** $\beta_0$ **? (1 point)**

Answer: 0.209295

**Based on the regression, what is the value of** $\beta_1$ **? (1 point)**

Answer: 0.079850

**Based on the regression, what is the impact of the treatment on the treatment group, relative to the control group, represented by** $\gamma$ **? (1 point)**

Answer: 0.079850

**Based on the regression, is the treatment effect statistically different from 0? (1 point)**

Answer: Yes, it is statistically significant.

### **Part 2. DiD with Fixed Effects**

**Question 4: Run a DiD model with the treatment group** $d_i$**, treatment period** $t$**, its interaction** $d_{it} \times t$, **and all other** **numerical variables in your data to estimate the treatment effect. What does the regression table look like? (1 point)**

```{r}
summary(mode2 <- lm(total_violent_crime~treatment + 
                      Population + White + Blk_AfAm + Asian + Hisp_Lat + 
                      Housing + Occupied + Vacant + Shape__Are + 
                      TLML + PRECTOT + SPEED, data = panel))
```

**Question 5: Based on the regression in Question 4, what is the** **treatment effect? Is the treatment effect statistically different from 0? (2 points)**

The treatment effect is 3.941e-02. Yes, it is statistically significant.

**Question 6: Run a DiD model the treatment group** $d_i$**, treatment period** $t$**, its interaction** $d_{it} \times t$, **and all other numerical, binary, and categorical variables. What does the regression table look like? (2 points)**

Instructions: A binary variable takes the form of 0 or 1. A categorical variable is a character with only a few unique values.

The efficient way to include categorical variables in a regression is to use the felm( ) function from the lfe package instead of the lm( ) function. The function has the format:

``` r
ModelName<- felm(OutcomeVariable ~ NumericalVariables + BinaryVariables | Categorical Variables, DataFrame)
```

The summary(ModelName) function is then used to print out the regression model.

```{r}
#| message: false
#install.packages("lfe")
library("lfe")
```

Write your model below and run it.

```{r}
Model3<- felm(total_violent_crime~treatment + 
                Population + White + Blk_AfAm + Asian + Hisp_Lat + 
                Housing + Occupied + Vacant + Shape__Are + 
                TLML + PRECTOT + SPEED + 
                treatment + total_violent_crime + total_property_crime | 
                Name + month + dayofweek + year, data = panel)
summary(Model3)
```

Feedback: Total property crime is a dependent variable as well and should not be included. Demographic variables are NaN because they are colinear with neighborhood fixed effects. Here is my suggested regression:

```{r}
Model3<- felm(total_violent_crime~treatment + 
                TLML + PRECTOT + SPEED | 
                Name + month + dayofweek + year, data = panel)
summary(Model3)
```

**Question 7: Based on the regression in Question 6, what is the** **treatment effect? Is the treatment effect statistically different from 0? (2 points)**

treatment affect: 2.221e-17, it is statistically different as the treatment effect has 3 stars.

**Question 8: Create more binary or categorical variables by multiplying binary variables or combining categorical variables together as you see fit. Explain what real-world factor each new variable represents. (2 points)**

Instructions: You created new variables by multiplying binary variables and combining categorical variables in asn-6-4-key.qmd Part 2. You explained what each variable represents in asn-6-4-key.qmd Part 1 Questions 4, 5, and 6.

```{r}
Model3<- felm(total_violent_crime~TLML + PRECTOT + SPEED + total_violent_crime + total_property_crime + SPEED:treatment| Name + year + month + dayofweek, data = panel)
summary(Model3)
```

Answer: The new SPEED:treatment variable represents the effect the wind speed has on the treatment variable, and if that's statistically significant.

**Question 9: Run a DiD model with the treatment group** $d_i$**, treatment period** $t$**, its interaction** $d_{it} \times t$, **and all other numerical, binary, and categorical variables that you have created in Question 8. What does the regression table look like? (2 points)**

```{r}
Model4<- felm(total_violent_crime~treatment + Population + White + Blk_AfAm + Asian + Hisp_Lat + Housing + Occupied + Vacant + Shape__Are + TLML + PRECTOT + SPEED + total_violent_crime + total_property_crime + SPEED:treatment| Name + month + dayofweek + year, data = panel)
summary(Model4)
```

**Question 10: Based on the regression in Question 9, what is the** **treatment effect? Is the treatment effect statistically different from 0? (2 points)**

Answer: The treatment effect is -9.522e-17, and the treatment is statistically different from 0 as it is marked with three stars.

**Question 11: Based on your intuition**, **which DiD regression was the best estimator of the treatment effect?** **Please explain why. (2 points)**

Answer: The DiD regression in question 10 is the best estimator as it includes the most variables and has the least margin of error.

### **Part 3 Thinking about causal inference**

**Question 12: A credible DiD analysis assumes that no other factor impacts the changes in the treatment group that have not been controlled. Based on your opinion, what factors can change the treated outcome that has not been accounted for? (10 points)**

Answer: The time of year, the population of each neighborhood, economic status of each neighborhood, gender.

**Question 13: Had you had more time, what additional variables (that you know exist) could you add to your data to take into account external factors that influence the treatment outcome? Please provide a link to each of the data sources. (4 points)**

Answer:

population: <https://bniajfi.org/2021/11/12/population-data-from-the-2020-census/>

economic status of each neighborhood: <https://bestneighborhood.org/household-income-baltimore-md/>

You have reached the end of this assignment. Commit and push your updates back into the GitHub repository. You can find the step-by-step tutorial [here.](https://drive.google.com/drive/u/0/folders/1eSMfNoV_ilp4OnID2Ql3YT2qXkhCS7Vm)
